### 2015-07-31 implementation-concretization.tex
New concretization textx

Figure~\ref{fig:concretization-time} shows the running time of concretization vs.
package DAG size in nodes.  We generated this plot by running {\tt concretize}
on all of Spack's 245 packages.  We see that for all but the 9 largest
packages, concretization takes less than 2 seconds on a Sandy Bridge machine, 
and less than 1 second on a Haswell MacBook Pro.  For larger DAGs, we begin
to see a quadratic trend, but even for 50 nodes or more, concretization
takes less than 6 seconds.
%
This performance is expected from a greedy, fixed-point method.
The algorithm repeats, evaluating conditions and adding dependencies and constraints, until
there are no more changes to be made.
%
%  Additionally, our current compiler concretization
% policy requires a search in the DAG; it prefers the compiler of the nearest ancestor
% if the current node's compiler is not set. 
%
More generally, concretization is an instance of the constraint
satisfaction problem, which is NP-complete. Our greedy solver is sufficient for the
packages we have built so far.  Its running time is insignificant
compared to the build time of most packages, and it only executes
once per build.
We do not expect to see software packages with thousands of dependencies in the near
future, and we do not expect concretization to become a bottleneck, even if we
use a full constraint solver.

>>>>

Figure~\ref{fig:concretization-time} shows the running time of concretization vs.
package DAG size in nodes.  We generated this plot by running {\tt concretize}
on all of Spack's 245 packages. We tested on three LLNL cluster front-end nodes:
a 2.3Ghz Intel Haswell node and a 2.6 GHz Intel Sandy Bridge node on our Linux
clusters, and the 3.6 GHz IBM Power7 front-end node of a Blue Gene/Q system.
Each point is the average of 10 trials.  We see that for all but the 10
largest packages, concretization takes less than 2 seconds on all machines tested.
For larger DAGs, we begin to see a quadratic trend, but even for 50 nodes or more, 
concretization takes less than 4 seconds on the Haswell machine and 9 seconds on the 
Power7.
%
This performance is expected from our greedy, fixed-point method, and
it is sufficient for the packages we have built so far.
Its running time is insignificant compared to the build time of
most packages, and it only executes
once per build. 
More generally, concretization is an instance of the constraint
satisfaction problem, which is NP-complete, so concretization could become
more costly.  However, we do not expect to see software packages with
thousands of dependencies in the near future, so we do not expect
concretization to become a bottleneck, even if we use a full constraint solver.
